\chapter{Adaptive Liquidity Provision via Reinforcement Learning}
\label{ch:rl_strategy}

\section{Introduction}
The inherent non-stationarity of financial markets, characterized by shifting volatility regimes, transient liquidity droughts, and non-linear microstructural dependencies, necessitates a quoting strategy that possesses high \textbf{adaptive capacity}. While the stochastic control framework derived in Chapter \ref{ch:microstructure} provides a theoretical benchmark, its reliance on simplifying assumptions (e.g., constant parameters, independent Poisson arrivals) limits its efficacy in fragmented, toxic environments.
 This chapter details the formulation of a \textbf{Deep Reinforcement Learning (DRL)} framework used to optimize market-making policies. We transform the continuous quoting problem into a discrete-time \textbf{Markov Decision Process (MDP)} and delineate the architectural and algorithmic innovations required to derive robust, hardware-accelerated execution policies.

\section{Markov Decision Process (MDP) Formulation}
We model the market-making task as a discrete-time MDP defined by the tuple $(\mathcal{S}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$. The agent (the market maker) interacts with the environment (the Limit Order Book) by observing the state $s_t$, taking an action $a_t$, and receiving a reward $r_t$ and a next state $s_{t+1}$.

\subsection{State Space Representation ($\mathcal{S}$)}
The state vector $s_t \in \mathcal{S}$ is a multidimensional summary of the instantaneous LOB state and the market maker's internal constraints. We define $s_t$ as:
\begin{equation}
    s_t = \left[ \rho_t, \sigma_t, q_t, \psi_{t-1}, \Delta S_t, \tau_t, \text{vpin}_t \right]
\end{equation}
Variables and features in the state vector:
\begin{itemize}
    \item $\rho_t$: The \textbf{Order Book Imbalance} at time $t$, representing the normalized volume difference between the best bid and ask ($L1$ imbalance).
    \item $\sigma_t$: The \textbf{rolling volatility estimate}, typically calculated over a sliding window of the last $N$ ticks to proxy current market risk.
    \item $q_t$: The \textbf{net inventory position} of the market maker, where $q_t \in [q_{min}, q_{max}]$.
    \item $\psi_{t-1}$: The \textbf{previous spread} utilized by the agent, allowing the model to incorporate "quoting inertia" and avoid jitter.
    \item $\Delta S_t$: The \textbf{mid-price innovation}, defined as $S_t - S_{t-1}$, capturing short-term momentum.
    \item $\tau_t$: The \textbf{time-to-horizon} parameter, $T-t$, which informs the agent about the remaining duration of the trading session.
    \item $\text{vpin}_t$: A proxy for \textbf{Order Flow Toxicity}, calculated using volume-synchronized probability of informed trading metrics.
\end{itemize}

\subsection{Action Space and Quoting Policy ($\mathcal{A}$)}
The agent's action $a_t \in \mathcal{A}$ determines the placement of bid and ask limit orders. To maintain deterministic execution within the FPGA fabric, we utilize a \textbf{discretized action space}. The agent selects a spread $\psi_t$ and a skew $\theta_t$:
\begin{equation}
    P_{ask}^{quote} = S_t + \frac{\psi_t}{2} + \theta_t, \quad P_{bid}^{quote} = S_t - \frac{\psi_t}{2} + \theta_t
\end{equation}
where:
\begin{itemize}
    \item $\psi_t$: The \textbf{bid-ask spread}, chosen from a finite set $\{ \psi_{min}, \dots, \psi_{max} \}$.
    \item $\theta_t$: The \textbf{price skew}, which shifts the entire quoting envelope to manage inventory or front-run anticipated price moves.
\end{itemize}
By discretizing these parameters into $M$ possible levels, we reduce the neural network's output layer size, facilitating low-latency hardware inference.

\subsection{Risk-Adjusted Reward Function ($\mathcal{R}$)}
The reward function is the most critical component of the DRL framework, as it encodes the market maker's utility. We define the instantaneous reward $R_t$ as:
\begin{equation}
    R_t = \Delta \text{PnL}_t - \eta \cdot \phi(q_t) - \zeta \cdot \mathcal{L}_{cost}(a_t)
\end{equation}
Variables in the reward function:
\begin{itemize}
    \item $\Delta \text{PnL}_t$: The \textbf{mark-to-market (MtM) profit and loss} change between $t-1$ and $t$. It includes captured spread from fills and the valuation change of the existing inventory $q_{t-1}$.
    \item $\eta$: The \textbf{inventory risk aversion coefficient}.
    \item $\phi(q_t)$: The \textbf{inventory penalty function}, often modeled as $\phi(q_t) = q_t^2$. This term penalizes the agent for holding large positions, forcing it to "lean" the book to neutralize risk.
    \item $\zeta$: The \textbf{transaction/execution cost coefficient}.
    \item $\mathcal{L}_{cost}(a_t)$: A penalty for \textbf{adverse selection} and quoting instability, discouraging the agent from placing orders that are immediately picked off by toxic flow.
\end{itemize}

\section{The Proximal Policy Optimization (PPO) Algorithm}
To derive the optimal policy $\pi_\theta(a|s)$, we employ the \textbf{Proximal Policy Optimization (PPO)} algorithm. PPO is an \textit{actor-critic} method that ensures stable learning by constraining the magnitude of policy updates.

\subsection{The Clipped Surrogate Objective}
The PPO agent maximizes the following objective function:
\begin{equation}
    L^{CLIP}(\theta) = \hat{\mathbf{E}}_t \left[ \min(w_t(\theta) \hat{A}_t, \text{clip}(w_t(\theta), 1-\epsilon, 1+\epsilon) \hat{A}_t) \right]
\end{equation}
Terms in the PPO objective:
\begin{itemize}
    \item $w_t(\theta) = \frac{\pi_\theta(a_t|s_t)}{\pi_{\theta_{old}}(a_t|s_t)}$: The \textbf{probability ratio} between the new and old policies.
    \item $\hat{A}_t$: The \textbf{estimated advantage} at time $t$, representing how much better the action $a_t$ is compared to the average action in state $s_t$.
    \item $\epsilon$: The \textbf{clipping hyperparameter} (typically $0.1$ or $0.2$), which prevents the new policy from deviating too far from the old one, ensuring monotonic improvement.
\end{itemize}

\subsection{Value Function Approximation and Advantage Estimation (GAE)}
The "Critic" network approximates the \textbf{Value Function} $V(s_t)$, which estimates the expected cumulative discounted reward from state $s_t$. To reduce variance in the gradient estimates, we use \textbf{Generalized Advantage Estimation (GAE)}:
\begin{equation}
    \hat{A}_t = \sum_{l=0}^{\infty} (\gamma \lambda)^l \delta_{t+l}
\end{equation}
where:
\begin{itemize}
    \item $\delta_t = R_t + \gamma V(s_{t+1}) - V(s_t)$: The \textbf{Temporal Difference (TD) error}.
    \item $\gamma$: The \textbf{discount factor}.
    \item $\lambda$: The \textbf{GAE smoothing parameter}, which balances the trade-off between bias and variance.
\end{itemize}

\section{Hardware-Centric Model Optimization}
Deploying deep neural networks (DNNs) on FPGA logic requires transforming the high-precision software model into a resource-efficient hardware core.

\subsection{Quantization-Aware Training (QAT) Mathematics}
Standard training occurs in FP32 (32-bit floating point). However, FPGA DSP48 slices are optimized for \textbf{Fixed-Point Arithmetic}. To prevent accuracy loss during conversion, we use \textbf{QAT}, which simulates the effects of low-precision quantization during the training process. The quantization function $Q(x, s, z)$ is:
\begin{equation}
    Q(x, s, z) = \text{clip}\left(\text{round}\left(\frac{x}{s} + z\right), q_{min}, q_{max}\right)
\end{equation}
Variables in quantization:
\begin{itemize}
    \item $x$: The high-precision weight or activation.
    \item $s$: The \textbf{scale factor}, $s = \frac{max - min}{q_{max} - q_{min}}$.
    \item $z$: The \textbf{zero-point offset}.
    \item $q_{min}, q_{max}$: The limits of the target integer range (e.g., $-128$ to $127$ for \text{INT8}).
\end{itemize}
By incorporating this "straight-through estimator" during backpropagation, the model learns weights that are inherently robust to the precision limits of the hardware.

\subsection{Weight Pruning and Sparsity}
To achieve sub-15ns inference, we apply \textbf{Iterative Magnitude Pruning (IMP)}. We remove the weights with the smallest absolute values, inducing \textbf{structured sparsity} in the network layers:
\begin{equation}
    W_{sparse} = W \odot M, \quad M_{i,j} = \begin{cases} 0 & \text{if } |W_{i,j}| < \text{threshold} \\ 1 & \text{otherwise} \end{cases}
\end{equation}
where $M$ is a binary mask. This sparsity allows the FPGA to skip redundant multiplications, drastically reducing power consumption and latency.

\section{Conclusion}
This chapter has detailed the transformation of the market-making task into a robust DRL framework. By synthesizing stochastic control theory with the PPO algorithm and hardware-centric QAT/pruning, we create a policy that is both microstructurally aware and execution-ready. The following chapters delineate the hardware-software architecture used to operationalize this policy at wire speed.
