\chapter*{Glossary of Financial and Technical Terms}
\addcontentsline{toc}{chapter}{Glossary}

\begin{description}
    \item[Adverse Selection:] A situation where a market participant trades with a counterparty who has superior information, often resulting in a loss for the less-informed party (e.g., a market maker being "picked off" by an informed trader).
    
    \item[Avellaneda--Stoikov Model:] A classical stochastic control model for market making that determines optimal bid and ask quotes based on inventory risk and market volatility.
    
    \item[Backtesting:] The process of testing a trading strategy using historical data to evaluate its performance before live deployment.
    
    \item[Best Bid and Offer (BBO):] The highest price a buyer is willing to pay (bid) and the lowest price a seller is willing to accept (offer) for a security at a given time.
    
    \item[Bid-Ask Spread:] The difference between the highest price a buyer will pay (bid) and the lowest price a seller will accept (ask). It represents a primary source of profit for market makers.
    
    \item[CAGR (Compound Annual Growth Rate):] The mean annual growth rate of an investment over a specified period of time longer than one year.
    
    \item[Capital Efficiency:] A measure of how effectively a firm or strategy uses its capital to generate returns or support trading activity.
    
    \item[Clearing:] The process of updating the accounts of the trading parties and arranging for the transfer of money and securities.
    
    \item[Co-location:] The practice of placing trading servers in the same data center as an exchange's matching engine to minimize network latency.
    
    \item[Direct Market Access (DMA):] A mechanism that allows trading firms to send orders directly to an exchange's order book without manual intervention by a broker.
    
    \item[Execution Venue:] A platform where trading of financial instruments occurs, such as a stock exchange (e.g., NASDAQ, NYSE) or an Alternative Trading System (ATS).
    
    \item[Fill:] The execution of a trading order. A "partial fill" occurs when only a portion of the requested quantity is executed.
    
    \item[Flash Crash:] An extremely rapid and deep drop in the price of one or more securities, often followed by a quick recovery, typically triggered by high-frequency trading algorithms.
    
    \item[Geometric Brownian Motion (GBM):] A continuous-time stochastic process often used to model stock prices, assuming a constant drift and volatility.
    
    \item[High-Frequency Trading (HFT):] A type of algorithmic trading characterized by high speeds, high turnover rates, and high order-to-trade ratios.
    
    \item[Inventory Risk:] The risk that a market maker will incur losses on their accumulated position due to adverse price movements.
    
    \item[Jitter:] The variation in time between the arrival of packets or the execution of tasks, often caused by operating system or network congestion.
    
    \item[Latency:] The time delay between a signal (e.g., a market data update) and the system's response (e.g., an order submission).
    
    \item[Limit Order:] An order to buy or sell a security at a specific price or better.
    
    \item[Liquidity:] The ease with which an asset can be bought or sold in the market without affecting its price.
    
    \item[Market Making:] A trading strategy where a participant provides liquidity by simultaneously quoting bid and ask prices, profiting from the spread.
    
    \item[Market Microstructure:] The study of the processes and outcomes of exchanging assets under a specific set of rules.
    
    \item[Market Order:] An order to buy or sell a security immediately at the best available current price.
    
    \item[Market Regime:] A period characterized by specific market conditions, such as high volatility or a strong trend.
    
    \item[Max Drawdown:] The maximum observed loss from a peak to a trough of a portfolio, before a new peak is attained.
    
    \item[Mid-Price:] The average of the current best bid and best ask prices.
    
    \item[Multicast Feed:] A method of sending market data where information is broadcast to multiple subscribers simultaneously over a network.
    
    \item[Notional Value:] The total value of a leveraged position's assets.
    
    \item[Order Book Imbalance (OBI):] A measure of the difference between the volume of buy orders and sell orders at the top of the order book, often used as a signal for short-term price movements.
    
    \item[Order Book:] A real-time, electronic list of buy and sell orders for a security or financial instrument, organized by price level.
    
    \item[P50 / P99 Latency:] Percentile measures of latency. P50 (median) is the latency below which 50\% of observations fall; P99 is the latency below which 99\% of observations fall.
    
    \item[Price Discovery:] The process by which the market determines the price of an asset through the interactions of buyers and sellers.
    
    \item[Price-Time Priority:] An execution rule used by exchanges where orders at the same price are filled in the order they were received.
    
    \item[Quoting Strategy:] The logic used by a market maker to determine the prices and quantities of their bid and ask quotes.
    
    \item[Relative Strength Index (RSI):] A momentum indicator that measures the magnitude of recent price changes to evaluate overbought or oversold conditions.
    
    \item[Reward Shaping:] A technique in reinforcement learning where the reward function is modified to guide the agent toward desired behaviors more efficiently.
    
    \item[Sharpe Ratio:] A measure of risk-adjusted return, calculated as the excess return per unit of volatility.
    
    \item[Slippage:] The difference between the expected price of a trade and the price at which the trade is actually executed.
    
    \item[Smart Order Router (SOR):] An automated system that determines the best venue and method to execute an order across multiple exchanges or liquidity pools.
    
    \item[Sortino Ratio:] A variation of the Sharpe Ratio that only penalizes "downside" volatility (returns below a user-specified target).
    
    \item[Stochastic Control:] A subfield of control theory that deals with systems whose state evolves over time in a probabilistic manner.
    
    \item[TCP Offload Engine (TOE):] A technology that offloads the processing of the TCP/IP stack from the CPU to a network adapter or FPGA.
    
    \item[Technical Indicator:] A heuristic or mathematical calculation based on the price, volume, or open interest of a security.
    
    \item[Temporal Market Dependencies:] Patterns or relationships in market data that persist over time.
    
    \item[Tick-to-Trade (T2T):] The latency from the receipt of a market data update to the submission of an order in response.
    
    \item[Toxic Flow:] Order flow that consistently moves the price against a market maker, typically originating from informed traders.
    
    \item[Zero-Copy Semantics:] A computer operation that eliminates the need for the CPU to copy data from one memory area to another, reducing latency.
\end{description}
